---
title: "Simple document"
output: github_document
---

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(haven)
library(shiny)
```



Let's import the `FAS_litters.csv` csv using a relative path.
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

litters_df =
  janitor::clean_names(litters_df)

litters_df 
```

Import the same dataset using an absolute path.
```{r}
litters_df_abs = 
  read_csv("data/FAS_litters.csv")

litters_df =
  janitor::clean_names(litters_df_abs)

litters_df_abs = 
 read_csv("~/Programming/P8105/data_wrangling_i/data/FAS_litters.csv")
```
Avoiding absolute paths are important because if you move the location, it is no longer reproducible. This is why it is better to keep an R project where everything is together.

Learning Assessment: Import pups dataset
```{r}
pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df =
  janitor::clean_names(pups_df)

pups_df_abs = 
  read_csv("~/Programming/P8105/data_wrangling_i/data/FAS_pups.csv")

pups_df_abs =
  janitor::clean_names(pups_df_abs)
```

## Look at data
```{r}
litters_df

head(litters_df)

tail(litters_df)
```

## You can use `view` function
```{r, eval=FALSE}
view(litters_df)
```

```{r}
str(litters_df)

skimr::skim(litters_df)
```

## Options in `read_*`


```{r}
litters_df = 
  read_csv(
    "data/FAS_litters.csv",
    skip = 10)

## useful when you have a data dictioanry in the first few rows of a dataset
```


Look at NA values

```{r}
litters_df = 
  read_csv(
    "data/FAS_litters.csv",
    na = c("NA", 19, "MISSING", "."))

# useful if you know that 999 is coded as "i dont know" for example 
litters_df = as_tibble(litters_df)
```

Column types 
```{r}
litters_df = read_csv(
  "data/FAS_litters.csv",
  col_types = cols(
    `GD0 weight` = col_character()
    # Specify other columns and their types here if needed
  )
)

```

## Other file types

 Import a xlsv file first.

```{r}
mlb_df = 
  read_excel("data/mlb11.xlsx")
```

import a sas datafile.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")
```


## Base R: do not do this...
```{r, eval = FALSE}
litters_df = 
  read.csv()

```



## Export Data

```{r}
litters_df_cleaned = 
  read_csv("data/FAS_litters.csv")

litters_df_cleaned = 
  janitor:: clean_names(litters_df_cleaned)

write_csv(litters_df_cleaned,"data/litters_cleaned.csv")
```


